---
title: Untitled
subtitle: A slightly longer title
format:
  360-analysis-html: default
author: James Goldie
date: last-modified
code-fold: true
---


```{r}
#| label: setup
library(tidyverse)
library(countrycode)
library(here)
```


```{r}
#| label: load-data
here("data", "raw") |>
  list.files(pattern = glob2rx("*.csv"), full.names = TRUE) |>
  map_dfr(read_csv) ->
all_data
```

```{r}
#| label: pre-natal
all_data |>
  filter(
    cause == "Neonatal preterm birth",
    metric == "Rate",
    age == "<28 days") |>
  select(location, rei, year, val, upper, lower) |>
  write_csv(here("data", "lt28days-allsex-preterm-pollution-deathrate100k.csv")) ->
preterms
```

```{r}
#| label: all-cause
all_data |>
  filter(
    is.na(cause)) |>
  select(location, rei, year, val, upper, lower) |>
  write_csv(here("data", "agestd-allsex-allcause-pollutiontemp-sev.csv")) ->
all_cause
```

Let's do the all-cause stuff wide by year and by measure, so we can show it in Datawrapper as a table.

:::{.callout-warning}
Datawrapper allows spans of columns to be joined as row-by-row line charts within a table, but if two of these line charts sit next to each other, they need a spacer column in between them.

Unfortunately, the title of that column shows up unless we hide it with CSS.

That, and the required column order, is why there's a bit of fiddliness here!
:::

```{r}
#| label: widen-allcause-yearsandrisks
all_cause |>
  select(-upper, -lower) |>
  mutate(rei = case_match(rei,
    "High temperature"                     ~ "hightemp",
    "Low temperature"                      ~ "lowtemp",
    "Ambient particulate matter pollution" ~ "pollution")) |>
  pivot_wider(id_cols = location, names_from = c(rei, year), values_from = val) |>
  # round figures off and add iso2 codes, then standardise country names
  mutate(
    across(where(is.numeric), ~ round(.x, 4)),
    iso2 = countrycode(location, "country.name", "iso2c"),
    country = countrycode(iso2, "iso2c", "country.name"),
    label = paste0(":", tolower(iso2), ": ", country),
    # datawrapper requires a couple of dummy 'spacer' columns for our
    # table-of-charts
    `<span style='display:none;'>S1</span>` = NA,
    `<span style='display:none;'>S2</span>` = NA) |>
  select(sort(tidyselect::peek_vars())) |>
  relocate(
    contains("<span style='display:none;'>S1</span>"),
    .after = starts_with("hightemp_")) |>
  relocate(
    contains("<span style='display:none;'>S2</span>"),
    .after = starts_with("lowtemp_")) |>
  relocate(label, .before = starts_with("hightemp_")) |>
  write_csv(
    here("data", "agestd-allsex-allcause-pollutiontemp-sev-wide.csv"),
    na = "") ->
all_cause_widest
```

And one more where we just widen by risk, not year:

```{r}
#| label: widen-allcause-byrisk
all_cause |>
  mutate(
    rei = case_match(rei,
      "High temperature"                     ~ "hightemp",
      "Low temperature"                      ~ "lowtemp",
      "Ambient particulate matter pollution" ~ "pollution"),
    iso2 = countrycode(location, "country.name", "iso2c"),
    country = countrycode(iso2, "iso2c", "country.name")) |>
  select(-location) |>
  pivot_wider(
    id_cols = c(country, year),
    names_from = rei,
    values_from = c(val, upper, lower)) |>
  write_csv(here("data", "agestd-allsex-allcause-pollutiontemp-sev-byrisk.csv")) ->
all_cause_byrisk
```